### 2013-04-05 doc/papers/cpc2013/paper.tex
First draft for CPC abstract submission
To be reviewed by Joël & Pierrick, and updated accordingly.

\documentclass{llncs}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\begin{document}
\title{Hybrid compiler-library approach to efficiently compile Numpy expressions}

\author{Serge Guelton\inst{1} \and Joël Falcou\inst{2} \and Pierrick Brunet\inst{3}}
\institute{ENS ULM, Paris, France \and LRI, University Paris Sud XI, France \and Télécom Bretagne, Plouzané, France}

\maketitle

\begin{abstract}

    The Numpy module for the Python language is growing in popularity among the
    scientific community as a scientific computing tool. It most noticeably
    provides a N-dimensional array object and a set of mathematical arrays
    using this structure that provide efficient computations thanks to the use
    of natively compiled implementation.

    Experiments shows that although more efficient than interpreted Python code,
    this module does not optimize complex array expressions and does not fully
    take advantage of multi-core machines with powerful vector instruction
    units. This paper examines the aggregation of complex Numpy expressions and
    their static compilation to efficient, hardware-aware native code to
    speedup mathematical computations.

    It shows that the combination of expression templates with a generic
    vectorisation library and OpenMP constructs can be used to turn the
    multiple sequential loops involved in complex Numpy expressions into a
    single parallel, vectorized loop that efficiently take advantage of the
    hardware capability.

    The proposed techniques are used to turn polymorphic Python functions to
    polymorphic C++ functions through template meta-programming. Specialization
    for given types is done during C++ template instantiation by the C++
    compiler.

    As the efficiency of the operator aggregation highly depends on the memory
    accesses over number of operations ratio, the paper also evaluates the
    relevancy of using forward substitution and eventually introducing
    redundant computations to increase computation intensity and hiding the
    memory latency implied by the multiple loads and stores to and from vector
    registers.

    The ideas presented in the paper are implemented in the Pythran compiler
    infrastructure using the Numeric Template Toolbox(NT$^2$). The static
    approach is compared with the Just-In-Time approach taken by the Numba
    compiler and the numexpr module, two tools used to dynamically optimize
    Numpy expressions.

\end{abstract}

%\bibliographystyle{splncs}
%\bibliography{biblio}

\end{document}

% vim:spell spelllang=en