### 2013-09-03 doc/papers/sc2013/paper.tex
Almost full revision of the paper
last benchmark part still needs to be written...

% to submit via
% https://www.easychair.org/account/signin.cgi?timeout=1;conf=pyhpc2013
% max 10 pages
\documentclass[conference]{IEEEtran}
%
\usepackage{listings}
\usepackage{url}
\usepackage{todo}
\usepackage{xcolor}
\newcommand\see{\emph{cf.\ }}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{shadows}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

\lstset{language=python,morekeywords={cimport, cdef, with},
basicstyle=\ttfamily, keywordstyle=\bf, stringstyle=\color{gray},
captionpos=b}

\begin{document}
% update listing numbering
\renewcommand{\thelstlisting}{\arabic{lstlisting}}
%
\title{Compiling Python Modules to Native Parallel Modules Using Pythran and
OpenMP Annotations}

\author{
    \IEEEauthorblockN{Serge Guelton\IEEEauthorrefmark{1}, \IEEEauthorrefmark{2}}
    \IEEEauthorblockN{Pierrick Brunet\IEEEauthorrefmark{2}}
    \IEEEauthorblockN{Mehdi Amini\IEEEauthorrefmark{3}}
    \newline
    \IEEEauthorblockA{\newline\IEEEauthorrefmark{1}
        {\'E}cole Normale Sup{\'e}rieure, D\'{e}partement d'Informatique, Paris, France}
    \IEEEauthorblockA{\newline\IEEEauthorrefmark{2}
    T{\'e}l{\'e}com Bretagne, Plouzan{\'e}, France}
    \IEEEauthorblockA{\newline\IEEEauthorrefmark{3}
    Silkan, Meudon la For\^et, France}
}

\maketitle

%
\begin{abstract}

    HPC users traditionally rely on low-level, compiled language such as C or
    FORTRAN to perform computation-intensive tasks. As a consequence, it is a
    common situation to have HPC application written in a high-level language
    such as Python, calling native routines for computation-intensive tasks. To
    improve development speed and reduce maintenance costs, several solutions
    such as Cython, Numba, Parakeet or Pythran, have appeared to make it
    possible to automatically or semi-automatically turn Python functions into
    native ones.

    One of the key point to match the performance of native applications is the
    ability to write parallel applications. This paper studies the addition of
    OpenMP directives, a popular model to describe the parallelism of
    C/C++/FORTRAN applications, to Pythran, an automatic compiler from a subset
    of Python to C++. It shows that scientific Python applications annotated
    with OpenMP directives can be turned by an automatic compiler into native
    applications that run within the same order of magnitude than
    manually-written ones.

\end{abstract}

%
\section{Introduction}

Since its birth in December 1989, the Python language~\cite{rossum97} has proved
to be useful in various domains, ranging from system administration to web
services, thanks to its dynamicity, expressiveness, rich ecosystem and
battery-included standard library. It is also getting widely used in scientific
computing~\cite{Oliphant2007}, mainly thanks to the SciPy~\cite{scipy} project
that provides a set of scientific and numeric tools, ranging from linear algebra
to ordinary differential equation solvers or generic algorithms.

It is now heading toward High Performance Computing(HPC), either as a glue
language used to bind several native libraries together, or to write whole
applications in Pure Python. However, the prohibitive performance overhead of
the language implied by its interpreted and highly dynamic nature prevents its
usage for the most performance - critical code sections, where native code is
generally used. The SciPy modules partly overcome this issue through the usage
of low-level routines written in C or Fortran and encapsulated in Python
\emph{native} modules. Moreover, its core data structure, the multi-dimensional
array~\cite{numpyarray2011}, has been designed so that the underlying data are
available to both native modules and the Python interpreter without conversion
cost. Yet when a new function not already part of an existing module is
required, one has to write it in C or FORTRAN.

To enable the user to write native functions without having to write the glue
code that turns Python object into native structures back and forth, the SciPy
package provides the \texttt{weave} module that makes it possible to bundle C
code snippet into Python code and compile then load them at runtime. This hybrid
approach, based on a mix of interpreted and native code, is getting widespread
in the Python landscape. It is also the recommended approach to write
application that use fine grain parallelism, has Python parallelism support
targets coarse-grain parallelism.

This paper presents the addition of OpenMP, a popular standard to turn
sequential C, C++ and FORTRAN application into parallel applications, to the
Pythran~\cite{pythran2013} compiler, a compiler that turn Python modules into
C++ meta-programs.  Section~\ref{sec:python-parallel} gives an overview of
existing approaches to compile Python functions or modules and shows a critical
lack of parallelism support, and emphasises on the need for a
backward-compatible approach.  Section~\ref{sec:python-static} presents the
Pythran compiler and the underlying runtime. Section~\ref{sec:python-openmp}
studies the validity of using OpenMP directives within the Python language in
order to add fine-grain parallelism support to Python in the context of Pythran.
The Pythran compiler is benchmarked on several scientific kernels, compared
to Cython, Numba and Parakeet in Section~\ref{sec:validation}.

%=========================================================
\section{Parallel Computations in Python}\label{sec:python-parallel}
%=========================================================

In the many cores area, turning to parallelism to balance the performance
limitations of scripting languages, as described in~\cite{choy05} in the context
of the MATLAB language, or in~\cite{mals07} in the context of the R language,
seems legitimate. In the context of Python, most approaches have focused on
fork-based parallelism.

\subsection{Python and Parallelism}

Parallel computations are supported by the Python standard library through the
\texttt{multiprocessing} module. It spawns several interpreters that can
communicate through Inter Process Communication(IPC), using Python built-in
object serialization. This approach is only viable for computation intensive
application, since the communication and synchronization overheads are much
greater than a light-weight processes approach.

The standard \texttt{threading} module also makes it possible to start
several light-weight threads within the same interpreter, but this lower-level approach is not
applicable to HPC, because of a specificity of CPython, the \emph{Global
Interpreter Lock}~\cite{gil2012}.\footnote{Other Python interpreters, such as
IronPython or Jython, do not have a GIL.} This lock ensures that only one thread
is active at a time in the interpreter. While it makes it possible to have
cooperative threads, say for a GUI, it does not take advantage of multiple
cores.  However, there are two notable exceptions: the GIL is released on I/O,
and the GIL does not prevent the use of threads inside native modules, where the
user has full control.

To illustrate the limitations of these two approaches, we used the CPU-bound
Buffon's needle algorithm to estimate the value of $\pi$. A sequential version
parallelized using the two approaches is illustrated in
Listing~\ref{lst:buffon-python}. When comparing their respective execution time
using 4 cores, it appears that the version that uses the \texttt{threading}
module runs $\times 1.46$ the execution time of the sequential version. GIL
contention actually \emph{increases} the execution time. The version that uses
\texttt{multiprocessing} provides a speedup of 3 over the sequential version
while a speedup of 4 would be expected for this kind of application.

\begin{lstlisting}[float, label={lst:buffon-python}, caption={Implementation of
sequential and parallel version of the Buffon algorithm in Python.}]
def buffon(darts):
  hits = 0
  for i in xrange (0, darts):
    x, y = random(), random()
    dist = sqrt(pow(x, 2) + pow(y, 2))
    if dist <= 1.0:
      hits += 1.0
  # hits / throws = 1/4 Pi
  pi = 4 * (hits / darts)
  return pi

from threading import Thread
from multiprocessing import Queue
def threaded_buffon(d):
  def work(darts, queue):
    queue.put(pi_estimate(darts))
  n = 4
  q = Queue()
  threads = [
    Thread(target=work, args=(d//n, q)),
    Thread(target=work, args=(d//n, q)),
    Thread(target=work, args=(d//n, q)),
    Thread(target=work, args=(d//n, q)),
  ]
  map(Thread.start, threads)
  map(Thread.join, threads)
  return sum(q.get() for _ in threads)/n

from multiprocessing import Pool
def multi_buffon(darts):
  p = Pool(4)
  return sum(p.map(buffon, [darts//4]*4))/4
\end{lstlisting}

There have also been several approaches to replace the GIL by Transactional
Memories~\cite{Riley2006,Tabba2010} but none of them made there way to the
mainstream interpreters.  As a consequence, Python developers need to write
multi-threaded native modules in order to fully benefit from multiple cores.
This leads to a kind of computations referred as \emph{hybrid computations}.

\subsection{Hybrid Computations}

In the context of interpreted languages, \cite{dongara2007} defines a computation
as \emph{hybrid} when part of the code is interpreted, and part of it is
executed natively.

It is now common for scripting language to have C bindings. To take advantage of
compiled code, and to overcome the GIL limitations, Python developers must write
parallel C/C++ functions and the associated boilerplate based on the Python C
API~\cite{pythoncapi}. Tools have been developed to relieve the user from
writing the encapsulating code, notably SWIG~\cite{swig2003} that relies on an
enhanced interface specification, or
\texttt{boost::python}~\cite{boostpython2007} that relies on C++ facilities to
guide translation.

An opposite approach consists in using the host language ---herein Python--- to
describe both parts of the system, and to let an automated tool perform the
translation to native code of a part of the application, generally the
computation-intensive one where parallelism has been expressed in some ways.
Thus, developers not familiar with lower level languages or not eager to invest
the additional development time can still benefit from a fair performance boost.
This approach is the subject of many studies that can be classified according to
their compatibility with the host language.

\subsubsection{Backward-Incompatible Approaches}

A constraining (from the performance point of view) aspect of the Python
language is its type system. It implies that each method call is resolved
dynamically, even a simple add operation. It comes at no surprise that many
approaches restrain the Python language to add a static typing overlay.  Also,
only two types of integers (64-bits integers and multi-precision integers) and
one type of floating point type (double-precision floats) are available in
Python, so using a type with the appropriate size may lead to significant
performance boost.

Cython~\cite{cython2010} is an hybrid Python/C dialect. It extends the Python
syntax with typing informations, calls to native functions from third party
libraries, and a limited set of parallelism constructs, such as the possibility
to define parallel loops, but no task parallelism. When possible, it unboxes
Python variable to improve performance. Without type annotation, the performance
improvement is not terrific, but given enough type annotation, Cython can
generates code that runs as fast as its C equivalent, while maintaining a syntax
close to Python.

PLW~\cite{dongara2007} and \texttt{Scipy.weave} both propose an other approach
that directly mixes Python with C, using raw strings to hold the C code. PLW
also supports parallel directives that are limited to parallel for loops.
PyCUDA and PyOpenCL~\cite{klockner2012} also target accelerators by mixing
Python with kernels embedded as raw strings containing accelerator code.

The main issue of these approaches is that they imply to modify, in a
backward-incompatible way, the original code. They require to learn a new
dialect, and the long-term preservation of this investment is not ensured.
Moreover, there is no falling back if the code were to be deployed in an
environment where the parallelizing tool/module is not available. For instance,
once a code has been ported to PyCUDA and transformed into CUDA code embedded in
Python strings, there is no way back and the developer has to maintain two
versions of the algorithm.

\subsubsection{Backward-Compatible Approaches}

Most backward-compatible approaches also require to modify the input program.
They do not extend the Python language, but restrict it. As a consequence, they
remain compatible with the original language and do not suffer from the
drawbacks of the previous approaches. They also benefit from existing tools
associated to the language.

Copperhead~\cite{copperhead2011} is a functional, data parallel language
embedded in Python. It uses $n$-uplets, NumPy arrays and lists as its core data
structure and prohibits usage of many control-flow operators such as loops,
enforcing the use of the \texttt{map}, \texttt{filter} or \texttt{reduce}
intrinsics to exhibit parallelism. But it can be efficiently compiled to either
CUDA or C++ with calls to the Thrust\footnote{\emph{cf}.\
\texttt{http://thrust.github.com/}} library. Python decorators are used to
identify hot-spots that are JIT-compiled to native code.

The numba~\footnote{\emph{cf}.\ \texttt{https://github.com/numba/numba}}
compiler uses additional type information to generate sequential LLVM bytecode.
Parakeet~\cite{parakeet2012} follows an approach similar to numba, that is
Python to LLVM bytecode translation, but limits its scope to the numpy package,
only supporting a small subset of the Python language. Additionally, it supports
implicit parallelism using an implicit mapping between numpy functions and a set
of parallel primitives including \texttt{map}s \texttt{scan}s and
\texttt{reduce}s. These two approaches use Just-In-Time(JIT) compilation and do
not suffer from the backward-incompatibility issues. When meeting an unsupported
construct, numba falls back to Python C-API calls, while parakeet raises an
exception.

Tools such as PyPy~\cite{pypy2009}, a Python interpreter with a tracing JIT, or
Shed~Skin~\cite{shedskin2006}, a Python to C++ compiler are also viable ways to
enhance Python performance. However Shed~Skin does not provide support for
fine-grained parallelism beyond what the standard library proposes. PyPy is
heading toward STM for parallelism support.

To be completely backward compatible, it must be possible to run the input code
in an environment that is not aware of the existence of the parallelizing
solution. This principle is not respected by parallel libraries, but code
annotations partially satisfy it. This is where parallel annotations shine: the
original code remains mostly compatible with a compiler that is not aware of the
annotations, while taking the annotation into account turns the sequential code
into parallel code.

This article proposes to combine OpenMP~\cite{openmp3.1} parallel annotations
with the Python subset supported by the Pythran compiler to make it possible to
write fine-grain parallel applications in Python, while being fully backward
compatible with both the Python language and the sequential algorithm. It means
that:

\begin{enumerate}

    \item any Pythran code can be run (sequentially), with no module dependency
        or code change, by any Python interpreter;

    \item parallelism is explicit and incrementally added to the original code
        through directives.

\end{enumerate}

%=========================================================
\section{Static Translation of Python Programs}\label{sec:python-static}
%=========================================================

Pythran~\cite{pythran2013} is a subset of Python designed for scientific
computing. It is implicitly statically typed and supports most Python constructs
except those that involve introspection (e.g.\ \texttt{getattr}) or runtime
compilation (e.g.\ \texttt{eval}). A few standard modules are supported in
addition to the core language (e.g.\ \texttt{math}, \texttt{random}). The
\texttt{numpy} module is currently only partially supported. User classes are
not supported. The associated compiler turns Pythran code, possibly annotated
with OpenMP and a few function type annotations, into C++ code. Its processing
is summarized in Figure~\ref{fig:pythran-compiler} and relies on type inference
to remove most dynamic behavior, combined with a high-level runtime library,
\texttt{pythonic++}, to allow a one-to-one mapping between Python and C++
constructs. The purpose of this section is not to dig into the internals of the
compiler, but rather to focus on the impact of the parallelism layer, especially
on the runtime library.

\begin{figure}

\centering
\begin{tikzpicture}[
file/.style={draw=black!50,fill=black!10,rectangle, drop shadow, align=center,
node distance=0.7cm},
    tool/.style={draw=black!50,fill=black!10,ellipse, align=center, node
distance=0.7cm}]
\node[file] (python) {\textbf{Python Module [\texttt{.py}]}};
\node[file] (omp)	   [xshift=-2em, left=of python] {\textbf{OpenMP}};
\node[tool] (pythran) [below=of python] {\textbf{Pythran}};
\node[file] (meta-cxx) [below=of pythran] {\textbf{C++} Meta Program};
\node[tool] (gxx) [yshift=-1.5em, below=of meta-cxx] {\textbf{g++}};
\node (empty) [xshift=-1em, left=of gxx] {};
\node[file] (pythonic) [left=of empty] {\textbf{\texttt{pythonic++}}};
\node[file] (annotation)	   [above=of pythonic] {\textbf{Type Info}};
\node[file] (boost) [below=of pythonic] {\textbf{\texttt{boost::python}}};
\node[file] (so) [yshift=-1.5em, below=of gxx] {\textbf{Native Module [\texttt{.so}]}};

\draw[very thick, ->] (python) -- (pythran);
\draw[very thick] (annotation) -| (empty.center);
\draw[very thick,dashed, ->] (omp) -- (python);
\draw[very thick, ->] (pythran) -- (meta-cxx);
\draw[very thick, ->] (meta-cxx) -- (gxx);
\draw[very thick] (boost) -| (empty.center);
\draw[very thick, ->] (pythonic) -- (gxx);
\draw[very thick, ->] (gxx) -- (so);
\end{tikzpicture}

\caption{Pythran compiler workflow.}
\label{fig:pythran-compiler}
\end{figure}

\subsection{Runtime Support}

Pythran runtime is based on the STL. The standard library of C++11, unlike its
C++03 version, is thread-safe, so no data dependencies are added to the original
code.

A critical point of the design of the \texttt{pythonic++} runtime is memory
management, the very same aspect that led to the use of a GIL in CPython.
Memory management is implemented in Shed Skin through the general-purpose Boehm
garbage collector~\cite{boehm1991}, and Cython forbids usage of Python-managed
objects inside parallel regions, thus making memory management explicit for
these parts.

Pythran handles the problem by refusing recursive types, which makes it possible
to solely rely on reference counting for memory management. It can be
implemented through the thread-safe shared pointer mechanism provided by the C++11
standard library.

Using shared references simplifies memory management, but the counterpart is an
extra atomic operation for each copy. As it limits parallelism, reducing the
number of copies becomes an important goal. The move semantic introduced in
C++11 avoids a few copies when working on temporary objects, but argument
passing still implies copies. To avoid a reference increment, one can pass
parameters by reference. Using an inter-procedural memory effect analysis not
presented in the paper, Pythran determines for each argument whether it must be
passed using reference or const reference to prevent this overhead.

\subsection{Directive Oblivious Translation}

During Python to C++ translation, Pythran adopts a blind strategy: it does not
understand the semantics of the annotations. Instead, it just splits each
annotation into a context-sensitive part ---the variable names--- and a
context-insensitive part ---the clauses---, and attaches them to the proper
instruction in the Abstract Syntax tree(AST).

The Pythran compiler ensures a bijective translation between Python instruction
and generated C++ instruction so that the OpenMP directive is regenerated on the
proper instruction. The same approach is used at the expression level, to be
compatible with the \texttt{if} clause.

At the AST level, it means that the Python AST is first reduced to a tree where
all nodes not available in C++ are transformed into a compatible representation.
For instance, list comprehension expressions are turned into function calls or
tuple unpacking is turned in multiple assignments. During this transformation
process, an expression is always transformed into a single expression and a
statement is always transformed into a single statement. Then this AST is
converted into a C++ AST meant to be pretty-printed.

\subsection{Transfer Costs}

When passing containers from Python to C back and forth, a copy of the whole
container is made to turn the type agnostic, non-contiguous Python data into
dense typed ones. This extra copy implies an extra cost that is not negligible:
Passing two lists of float from Python to C++ requires as much as half the time
to compute the dot product of the same lists directly in Python. Following
Amdahl's law, this copy overhead greatly hinders the benefits of
parallelization, and is a well known issue.

The traditional solution is to use a native type that exposes a Python interface
and has a constant translation cost. NumPy's \texttt{ndarray} is a typical
implementation of this concept and is the keystone of the Scipy and NumPy
packages. Basically, a NumPy \texttt{ndarray} is a raw C pointer that is exposed
at the Python level. As a tool for scientific computing, Pythran supports such a
structure, implemented as a class wrapping a raw pointer.

%=========================================================
\section{OpenMP Semantical Adaptations}\label{sec:python-openmp}
%=========================================================

OpenMP is a standard API for parallel programming for Fortran, C and C++. It
consists of a set of parallelizing directives and a few runtime library calls.
If OpenMP is not activated, the directives are ignored, thus enabling
incremental parallelization of the original source code while keeping the
original code structure. The languages targeted by OpenMP are statically
compiled. This section studies the semantical adaptation required to use the
same API for a scripting language such as Python.
Listing~\ref{lst:motivating-example} introduces a motivating example through the
computation of $\pi$, featuring a parallel reduction. The C equivalent with
OpenMP directives is given in Listing~\ref{lst:motivating-example-c} for
reference.

\begin{lstlisting}[float, label={lst:motivating-example}, caption={Motivating
example: computing $\pi$ in Python.}]
def pi(nsteps):
  sum, step = 0., 1. / nsteps
  for i in range(nsteps):
      x = (i - 0.5) * step
      sum += 4. / (1. + x**2)
  return step * sum
\end{lstlisting}

\begin{lstlisting}[float, label={lst:motivating-example-c}, caption={Motivating
example: computing $\pi$ in C with OpenMP.}]
double pi(size_t nsteps) {
  double sum = 0., step = 1. / nsteps;
  #pragma omp parallel for reduction(+:sum)
  for(size_t i = 0; i < nsteps ; ++i)
  {
      double x = (i - 0.5) * step;
      sum += 4. / (1. + x * x);
  }
  return step * sum;
}
\end{lstlisting}

\subsection{Directives}

OpenMP directives are held by C/C++ \texttt{\#pragma}, or by Fortran comments. Most of them
apply to structured blocks, represented by an instruction in C/C++ and delimited
by comments in Fortran. A few directives (e.g.\ \texttt{threadprivate}) are not
attached to a specific instruction.

While Python has a decorator mechanism\footnote{\see PEP 318,
\url{http://www.python.org/dev/peps/pep-0318/} for a more detail explanation of
Python decorators.}, it only applies to functions, methods and classes and does
not make it possible to attach decorations to other statements.
PLW~\cite{dongara2007}, uses string instructions to hold such decorations. As
Python does not have anonymous block,~\footnote{\see PEP340,
    \url{http://www.python.org/dev/peps/pep-0340/} for a discussion concerning
anonymous block support in Python.} one has to create a dummy \texttt{if 1:}
instruction to apply an annotation to a whole block. Pythran uses comments to
hold OpenMP directives, and internally stores them as string instructions.
Alternatively, the syntax \texttt{if 'my annotation':} is also supported.

Many OpenMP annotations are parametrized by clauses that list variables,
specifying their behavior with respect to parallel regions, e.g.\
\texttt{private}, \texttt{shared}, \texttt{copyin}. They can only refer to
variables that have already been declared. However, there is no variable
declaration in Python, and all variables assigned in a function have the
function scope, called local scope and available through the \texttt{local()}
built-in. As a consequence, all variables that are referenced in a function are
considered when building such variable lists: there is no concept of variable
local to a block. Using the \texttt{default(none)} clause is possible to ensure
no variable gets forgotten when building such lists.

The \texttt{reduction(\emph{operator}: \emph{list})} directive is used to
characterize some data dependencies when performing a parallel reduction. The
list of supported operators depend on the input and backend languages: Python
has \texttt{min}/\texttt{max} operators but C/C++ does not. C/C++ have
\texttt{\&\&} or \texttt{||} while Python does not.\footnote{Although similar,
the \texttt{and} and \texttt{or} keywords have a special meaning in Python.} The
latest OpenMP specifications~\cite{openmp4} describe a way to declare
user-defined reduction but is not implemented in any compiler yet.

The annotated motivating example is given in
Listing~\ref{lst:motivating-example-annotated-v0}. Note that unlike in C,
\texttt{x} has to be listed in a \texttt{private} clause.

\begin{lstlisting}[float, label={lst:motivating-example-annotated-v0}, caption={Motivating
example: computing $\pi$ in Python with OpenMP.}]
def pi(nsteps):
  sum, step = 0., 1. / nsteps
#omp parallel for reduction(+:sum) private(x)
  for i in range(nsteps):
      x = (i - 0.5) * step
      sum += 4. / (1. + x**2)
  return step * sum
\end{lstlisting}

\subsection{Automatic Scoping Computation}

Using function scope for all variables may lead to very long \texttt{private}
clauses. This section describes an algorithm to automatically limit the scope of
a variable declaration so that it is can be automatically marked as private, in
a similar manner to C/C++

The whole idea is to compute the minimal nesting depth of a variable usage. To
do so, a \emph{nesting depth} is first associated to each expression. It
lexically corresponds to the indentation level of its instruction, to the
notable exception of the iterator declaration in a for loop, that holds a
nesting depth equals to 1 plus the loop declaration indentation level.

Tracking the nesting depth of all variable use, for instance using def-use
chains, makes it possible to build a list of nesting depths, one per variable
reference. Computing the min of this list yields the scope of the variable.

For instance, in the motivating example, the \texttt{nsteps} variable is first
defined as argument, then read twice, which gives a def-use chain of
\{DEF, USE, USE\}. The associated nesting depths is
\{0, 1, 1\} so the variable scope is 0. \texttt{i} is defined as a loop iterator then
used in the loop body, which gives a chain of \{DEF, USE\}, with nesting
depths of \{2, 2\} and a scope of 2. Likewise \texttt{x} has a chain
of \{D, U\}, nesting depths of \{2, 2\} and a scope of 2, which means it can be
automatically declared local to the loop, thus not needing to be listed as
\texttt{private}.

To be able to list a variable marked as local, for instance in a
\texttt{lastprivate} clause, the directive accesses are taken into account in
the def-use computations. For instance in
Listing~\ref{lst:motivating-example-annotated-v0}, the chain for \texttt{x} is
\{DEF, DEF, USE\} and the associated nested depths are \{1, 2, 2\} so its scope
is 1 and the variable is not declared local to the loop.

Thanks to this automatic scoping computation, it is possible to write shorter
OpenMP directive, as illustrated in
Listing~\ref{lst:motivating-example-annotated-v1}.

\begin{lstlisting}[float, label={lst:motivating-example-annotated-v1}, caption={Motivating
example: computing $\pi$ in Python with OpenMP and automatic scoping.}]
def pi(nsteps):
  sum, step = 0., 1. / nsteps
  #omp parallel for reduction(+:sum)
  for i in range(nsteps):
    x = (i - 0.5) * step
    sum += 4. / (1. + x**2)
  return step * sum
\end{lstlisting}

\subsubsection{Parallel For and Iterators}

The core directive to handle data parallelism is the \texttt{for} directive that
distributes the iteration space of the associated loop among the existing
threads. To be compatible with OpenMP, the loop iteration space must be
described by a random access iterator with a total order. Integers used as loop
indices in Fortran and C satisfy this conditions, as well as C++ iterators with
the \texttt{random\_access\_tag} trait. But a Python iterator only advances by a
step of one until it is exhausted, in which case it raises an exception: it
behaves as a forward iterator. To be compatible with OpenMP, these iterators
must be turned into random access iterators.

The extension of Python iterator to random access iterators is direct for the
standard containers: \texttt{list}\footnote{Python \texttt{list}s behave as C++
\texttt{vector}s.}, \texttt{set} or \texttt{dict}. Other iterators require more
care.

Generators, Python objects that behave like iterators, are commonly used in
Python. The simplest one, \texttt{xrange(\emph{start}, \emph{stop},
\emph{step})}, successively yields value starting from \texttt{\emph{start}} to
\texttt{\emph{stop}} by a step of \texttt{\emph{step}}. It is easily extended to
support random access, but it generally does not make sense to use a generator
as a loop iterator, as the relation between two random states of the iterator
may be of an arbitrary complexity.

Generator expressions are generators whose content is built from an other
iterator. For instance \texttt{(x*x for x in l)} successively yields the square
of each element in \texttt{l}. They behave like adaptors: they apply a
particular expression on each value of the iterator. It is also the case of the
\texttt{enumerate} builtin, that yields each element of the enumerated iterator
associated with its index. These generators are random access iterators only if their
input iterator is a random iterator itself.

Finally, if the iterator yields a tuple, it is possible to unpack it inside the
\texttt{for} construct, as shown in Listing~\ref{lst:complex-omp-for}. In that case
all the unpacked variables are considered as iterators, especially with respect
to default privatization rules: in the given example, \texttt{i} and \texttt{v}
are private, and the parallel iteration is valid because the input of
\texttt{enumerate} is a list, which allows random access iteration.

\begin{lstlisting}[float, label={lst:complex-omp-for}, caption={Parallel
    loop in Pythran with tuple unpacking.}]
'omp parallel for'
for i, v in enumerate([2, 3, 5, 7, 11]):
    print i, ':', v
\end{lstlisting}

Pythran's runtime is aware of these three kind of iterators and supports
parallel iteration over random access iterators and iterators adapting random
ones.

\subsection{OpenMP Runtime Library}

OpenMP provides a small runtime library that, for instance, makes it possible to
retrieve the active thread id. All the functions are declared in the
\texttt{<omp.h>} header, and have a default behavior when OpenMP is not
activated.

Providing a binding to these library in Python as an \texttt{omp} module does
not raise particular problems, as the signature of these functions only involves
integers passed as parameters or return values, except for the mutex
manipulation. In that case an opaque type is used to represent the native type.

The \texttt{\_OPENMP} macro definition is always provided when OpenMP is
activated, and can be used to detect when OpenMP is not available. Python does
not have a preprocessor, but it is possible to catch the import exception if the
\texttt{omp} module is not found. Listing~\ref{lst:sample-api-usage} showcase
such a call using an example converted from the OpenMP validation suite
presented in next section. The code starts a parallel section using the
\texttt{parallel} directive, spawning several threads. Then each started thread
increases the \texttt{nthreads} variable that has unspecified visibility, thus
\texttt{shared}. The incrementation is protected by a guard through the
\texttt{critical} directive. Then one of the thread retrieves the number of
active thread in the parallel region through the \texttt{get\_num\_threads}
function from the \texttt{omp} module. The function should always return
\texttt{True}.

\begin{lstlisting}[float,label={lst:sample-api-usage},caption={Example of
        OpenMP API usage from Python.}]
import omp
def omp_get_num_threads():
  nthreads, nthreads_lib = 0, -1
  #omp parallel
  if 1
    #omp critical
    nthreads += 1
    #omp single
    nthreads_lib = omp.get_num_threads()
  return nthreads == nthreads_lib
\end{lstlisting}

\subsubsection{Interaction With Pythran Runtime}

Pythran already uses OpenMP to parallelize some Python functions. For instance
the \texttt{sum} function from the \texttt{\_\_builtin\_\_} is implemented using
the OpenMP \texttt{reduction} clause. When it can prove that its first argument
is a \emph{pure} function, it can also call a parallel version of the
\texttt{map} function that internally spawns a parallel region. This leads to a
well known situation with OpenMP, where too many threads may be spawned: nested
parallel regions. OpenMP 4 provides several ways to handle this situation:

\begin{enumerate}

    \item Disabling nested parallelism, through the \texttt{OMP\_NESTED}
        environment variable or the \texttt{omp\_set\_nested} routine.

    \item Setting the maximum number of active levels, through the
        \texttt{OMP\_MAX\_ACTIVE\_LEVELS} environment variable or the
        \texttt{omp\_set\_max\_active\_levels} routine.

    \item Conditionally activating regions or conditionally setting the number
        of threads in a parallel region using the \texttt{omp\_in\_parallel}

\end{enumerate}

The decision is left to the user, depending on its application and OpenMP
implementation.

\subsection{Validation}

A validation suite for OpenMP is proposed in~\cite{wang2012} for C and Fortran.
We ported it to Python, and also extended it to validate the corner cases
specific to Python described in this section. A typical test case is given in
Listing~\ref{lst:openmp-validation}.

\begin{lstlisting}[float,label={lst:openmp-validation},caption={Example of Python
    OpenMP validation test case.}]
def omp_parallel_for_if(loop_count):
  import omp
  using = sum = 0
  'omp parallel for if(using == 1)'
  for i in range(loop_count + 1):
    num_threads = omp.get_num_threads()
    sum += i
  known_sum = (loop_count *
              (loop_count + 1)) / 2
  return known_sum == sum and num_threads == 1
\end{lstlisting}

We have used the Pythran tool described in the following section to turn each
Python test function into a C++ function with the same directives and runtime
calls. Apart from the \texttt{threadprivate} directive and the
\texttt{collapse(\emph{n})} clause, all tests were successful.
\texttt{threadprivate} directives were held by global variables not supported in
Pythran yet~; and the C++ code generated by Pythran does not preserve the perfect
loop nesting required by the \texttt{collapse} clause.

%=========================================================
\section{Validation}\label{sec:validation}
%=========================================================

In addition to the formal validation done using the OpenMP validation suite
presented in Section~\ref{sec:python-openmp}, the performance of parallel module
generated by Pythran from Python module with explicit parallelization using
OpenMP directives has been studied on several situations: first on the
motivating example, then on a synthetic geomatic application. Python and Cython
versions are then compared and finally the effect of OpenMP directives on Pythran
generated code are illustrated in the context of the Python Benchmarks
suite~\footnote{\url{https://github.com/numfocus/python-benchmarks}}.

The target machine has  8 AMD Opteron 6176 SE cores and 8 GO of RAM. It is
running a Linux kernel 3.10-2-amd64, x86\_64 GNU/Linux. The C++ compiler used is
\texttt{g++} version 4.8.1 and all applicationsa re compiled using the
\texttt{-Ofast} flag. Pythran generated programs are linked with Boost.Python
version 1.54. The Pythran version is extracted from the \texttt{sc2013} branch
of the git repository \url{https://github.com/serge-sans-paille/pythran}. Python
programs are run using the CPython implementation, version 2.7.5. Cython
programs are generated using version 0.19.1. Numba and Parakeet are installed
from the \texttt{master} branch of their respective repositories.

\subsection{Motivating Example}

To validate the approach proposed in this paper, let first consider the
motivating example. Figure~\ref{fig:motivating-example-scale} shows how the C
and Python version of the program scale when changing the number of core used.
As a reference, the original Python code runs 41 times slower than the
equivalent C code. As the program is not memory bound and exhibit trivial
parallelism, the C + OpenMP version scales linearly. The Pythran + OpenMP
version follows the same pattern.

\begin{figure}

    \caption{Comparison of the scaling of $\pi$ computation function for C and
    Python with OpenMP, depending on the number of active threads.}
    \label{fig:motivating-example-scale}

    \includegraphics[width=.5\textwidth]{pi_omp_bench.pdf}

\end{figure}

\subsection{Hyantes: a Geomatic Application}

To get closer to a realistic example, we have considered the code of a small
geomatics application, Hyantes. Starting from the C code, we successively
parallelized it with OpenMP, turned it into Python with a Pythran compatible
kernel and parallelized the Python version using the approach described in this
paper. We also measure the SLOC of the two versions.  Table~\ref{tbl:hyantes}
summarizes the results of the experimentation. It not only emphasises the
use of a high level language prototyping, but also shows that it is possible to
turn this prototype into reasonably efficient code through Pythran. It is also
possible to prototype the parallel version while remaining at the Python level.

\begin{table}

    \caption{Comparison of several versions of the Hyantes prototype.}
    \label{tbl:hyantes}

    \centering
    \begin{tabular}{|l|r|}
        \hline
        Language & SLOC  \\
        \hline
        C       & 102   \\
        Python  & 30    \\
        Pythran & 30     \\
        Pythran+OMP    & 31  \\
        \hline
    \end{tabular}

\end{table}

The Hyantes programs scales relatively well. Figure~\ref{fig:hyantes-speedup} shows
the relative speedup of the application when increasing the number of OpenMP
threads using up to 8 cores.

\begin{figure}

    \caption{Comparison of the scaling of the hyantes application for C and
    Python with OpenMP, depending on the number of active threads.}
    \label{fig:hyantes-speedup}

    \includegraphics[width=.5\textwidth]{hyantes_omp_bench.pdf}

\end{figure}

\subsection{Comparison with Cython}

We then compare Pythran with Cython. The approaches share some similarities:
they both generate code written in a lower level language, with OpenMP
directives. However, the inputs differ as Cython requires more typing
information than Pythran to generate efficient code, and Cython is not
backward-compatible with Python. They translate explicit fine-grained
parallelism through code annotations or specific Python constructs,
respectively.

Pythran exposes the full OpenMP interface to the user, thus enabling both data
and task parallelism, as described in Section~\ref{sec:python-openmp}. It is
not the case in Cython:
%
\begin{itemize}

    \item Only loops can be made parallel, using a new \texttt{prange} generator.

    \item Reductions and variable privacy are inferred, but it chokes on
        reduction on private variables.

    \item The user is responsible from releasing the GIL inside parallel
        regions.

    \item It is impossible to use a function imported from a Python module in a
        parallel region, but it is still possible to use native C functions.

\end{itemize}
%
Listings~\ref{lst:cython-sample} and~\ref{lst:pythran-sample} illustrate the
difference between the two and illustrates the intrusive behavior of Cython.

\begin{figure}

    \begin{lstlisting}[label={lst:cython-sample}, caption={Cython implementation
    of a parallel reduction.}]
from libc.math cimport sqrt
from cython.parallel import parallel, prange
def sum_sqrt(double r):
    cdef int i
    with nogil, parallel():
        for i in prange(10000000):
            r += sqrt(i)
    return r
    \end{lstlisting}
%
    \begin{lstlisting}[label={lst:pythran-sample}, caption={Pythran implementation
    of a parallel reduction.}]
#pythran export sum_sqrt(float)
import math
def sum_sqrt(r):
    #omp parallel for reduction(+:r)
    for i in xrange(10000000):
        r += math.sqrt(i)
    return r
    \end{lstlisting}
\end{figure}

The performance of the two approaches is shown in
Figure~\ref{fig:cython-pythran}.  The benchmarked codes are typical
mathematical, image-processing or linear algebra kernels. All these kernels have
been written in Cython and Python ---compatible with Pythran--- and annotated
through the mechanism of each language to exhibit parallelism, then compiled
into native code.  Their execution time when called from the Python interpreter
is measured through the \texttt{timeit} module. All results are normalized
against Cython sequential execution time. They show that while handling code at
a higher level than Cython, Pythran achieves comparable results.

The source codes used for these benchmarks are available on the Pythran
repository.\footnote{\see the \texttt{sc2013} branch of the git repository.}

\begin{figure}[ht]
    \includegraphics[width=.5\textwidth]{cython}
    \caption{Comparison of Cython and Pythran generated code performance.}
    \label{fig:cython-pythran}
\end{figure}

\subsection{Python Benchmarks Test Bed}

To Be FINISHED

\begin{figure}[ht]
    \includegraphics[width=.5\textwidth]{growcut}
    \caption{Comparison of Various python Compilers on the Growcut benchmark}
    \label{fig:pb-growcut}
\end{figure}

\begin{figure}[ht]
    \includegraphics[width=.5\textwidth]{julia}
    \caption{Comparison of Various python Compilers on the Julia benchmark}
    \label{fig:pb-julia}
\end{figure}

\begin{figure}[ht]
    \includegraphics[width=.5\textwidth]{rosen_der}
    \caption{Comparison of Various python Compilers on the rosen\_der benchmark}
    \label{fig:pb-rosen}
\end{figure}

\begin{figure}[ht]
    \includegraphics[width=.5\textwidth]{pairwise}
    \caption{Comparison of Various python Compilers on the pairwise benchmark}
    \label{fig:pb-pairwise}
\end{figure}
growcut: $\times 2.65$
julia: $\times 3.95$
rosen\_der: $\times 1.78$
pairwise: $\times 4.27$

%=========================================================
\section{Conclusion and Future Work}
%=========================================================

This paper studies the addition of OpenMP annotations to Pythran. It shows that
providing minor semantic adaptations, these annotations make it possible for
Python code to benefit from multi-cores while retaining backward compatibility
and without worrying about the Global Interpreter Lock.

To achieve this goal, a translator from a subset of the Python language to C++
has been designed. This translator turns regular Python modules annotated with
OpenMP directives and a few type annotations into native parallel module. The
input module remains compatible with the standard interpreter and the underlying
runtime is compatible with parallel constructs.

The approach is compared with Cython, an extension of the Python language used
to generate native module with an hybrid Python-C syntax that also provides
means to exhibit fine grained parallelism. It shows that retaining Python
compatibility does not prevent the achievement of comparable performances.
Comparisons to Numba and Parakeet shows the benefit of using parallelism on the
benchmarks from the Python Benchmark project.

Future work will focus on the extension of the approach to OpenMP 4 and the
\texttt{target} clause that should make it possible to target MIC from Python.
The \texttt{simd} clause also brings interesting vectorization capabilities that
can be exposed at the python level.  There are also several implicit vectorization
opportunities in Python, especially in the list comprehension construction, that
needs to be explored.

%=========================================================
\section*{Acknowledgment}
%=========================================================

This work has received fundings from \textsc{silkan} and the French ANR through
the CARP project. The authors thank Adrien \textsc{Merlini} and Alan
\textsc{Raynaud} for their work on the runtime, Albert \textsc{Cohen},
B{\'e}atrice \textsc{Creusillet}, Fabien \textsc{Dagnat}, Fran{\,c}ois
\textsc{Irigoin} and Ronan \textsc{Keryell} for their valuable reviews.

\bibliographystyle{IEEEtranS}
\bibliography{biblio}

\end{document}
% vim:spell spelllang=en
% vim:tw=80