### 2013-09-27 doc/papers/cc2014/paper.tex
Setup format for cc2014 submission

\documentclass{llncs}

\usepackage{listings}
\usepackage{url}
\usepackage{todo}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{shadows}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

\bibliographystyle{splncs}
\lstset{language=python,morekeywords={cimport, cdef, with},
basicstyle=\ttfamily, keywordstyle=\bf, stringstyle=\color{gray},
captionpos=b}

\title{Pythran: Turning Python Modules into Optimized C++ Meta-Programs}

\author{Serge Guelton\inst{1,2} \and Pierrick Brunet\inst{2} \and Mehdi Amini\inst{3}}

\institute{École Normale Supérieure, Département d'Informatique, Paris, France
\and
Télécom Bretagne, Plouzané, France
\and
SILKAN Inc., Los-Altos, USA
}
\begin{document}

\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}
...
\section{Introduction}

\bibliography{biblio}

\end{document}