### 2014-01-27 doc/papers/CGO2014/paper.tex
Add CGO 2014 draft

\documentclass[10pt, onecolumn, preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{listings}
\usepackage{hyperref}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\lstset{captionpos=b, float, language=python}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{shadows}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

\providecommand{\boostsimd}{\textsc{Boost.SIMD}}
\providecommand{\cpp}[1][~]{\textsc{C++}#1}
\providecommand{\ie}[1][~]{\textit{i.e.}#1}
\providecommand{\eg}[1][~]{\textit{e.g.#1}}

\begin{document}

\title{Pythran: Benchmarking parallel Scientific Python Programs}

\authorinfo{Serge Guelton}
           {T{\'e}l{\'e}com Bretagne}
           {serge.guelton@telecom-bretagne.eu}
\authorinfo{Pierrick Brunet}
           {INRIA/MOAIS}
           {pierrick.brunet@inria.fr}
\authorinfo{Mehdi Amini}
           {SILKAN}
           {mehdi.amini@silkan.com}

\maketitle

\begin{abstract}

    Turning Python module in native one is a common practice in the Python
    eco-system to improve performance of computation intensive kernel for
    scientific modules. Many Python compiler are born in the Python community
    such as the well known Cython trying bring in Python the holy C
    performances. To reach this goal, optimisations should be done
    using as much as possible computer capabilities instead of doing a
    straightforward conversion. Such optimisations should take into account
    multi-core architecture and vectorization available in common
    platforms as numpy - the reference scientific python module - do an
    intensive use of such functions. Thanks to these optimisation,
    Python module should be able to achieve higher performances.

    Pythran is an open source static compiler that turns modules written
    in a subset of Python into native ones~\cite{pythran}. It already apply
    some optimisations like constant-folding for functions or partial forward
    substitution. This paper introduce integration of vectorization and
    parallelism on numpy array based kernel. Pythran already provides
    possibility to add OpenMP annotation in Python program for explicit
    parallelism~\cite{pythran-compas} and now, it aims to add it through its
    C++ runtime library using the Numeric Template Toolbox (NT2) for vectorization
    and OpenMP for parallelism.

    This approach have been validated though benchmark between Pythran with
    vectorization and parallelism and Pythran without it. It also compares these
    measures with others Python compiler like Numba and Coperhead.

\end{abstract}

\keywords
static compilation, parallelization, Python, C++, SIMD

%%
%%
\section{Introduction}

L'optimisation de performances pour les programmes Python n'est pas récent
puisque les premiers compilateur de pseudo language proche de Python comme Pyrex
ont plus de 10 ans maintenant. Certains projets sont aujourd'hui arrété
alors que d'autre naissent pour apporter toujours plus de performances.

%Cython
%De nombreux compilateur s'intéresse à la vectorization. automatique. Dans la
%communauté Python, Cython prévois des améliorations à ce sujet bien que rien ne
%soit encore implémenté
%\footnote{\url{https://github.com/cython/cython/wiki/enhancements-simd}}.
%Pour le multi-core, Cython ajoute une librairy mais ne fait rien
%automatiquement (donc c'est des annotations cf papier renpar)
%
%Nuitka
%je ne sais pas ce qu'ils font. Rien sur le web, il faudrait regarder le code
%généré pour savoir... (Dans le code, aucune trace de parallelism SIMD/OpenMP)
%
%Numba
%ne peut faire du parallelism que si il ne repose pas du tout sur Python
%(qu'il compile tout le code) : \url{http://www.phi-node.com/2013/01/just-in-time-compilers-for-number.html}
%Ils ont un prange comme Cython pour avoir du parallélisme explicite et ils ont
%une library de threading a eux...
%Pour le SIMD, ils font de la fusion d'operateur SI l'utilisateur le demande et
%specifie correctement les dimensions/profondeur ou on applique la fusion:
%\url{http://numba.pydata.org/numba-doc/0.13.4/ufuncs.html}
%Ils font du déroulage partiel de boucle et du vectoriel ensuite mais on a pas
%beaucoup d'info je trouve... (J'ai trouvé ça dans le code avec du code :
%oldnumba/minivect/specializers.py:438: rhs = b.vector\_load(data\_pointer, self.sp.vector\_size)
%Ils prévoient aussi un support openmp apparement : "OpenMP is not yet implemented, only process the 'else' directives. llvm\_codegen.py l.168"
%Ils ont un bon support cuda par contre : \url{https://developer.nvidia.com/how-to-cuda-python}
%
%Numexpr
%fait de la composition d'opérateur mais pas d'utilisation de vectorization
%Ils ont un support de parallelisation multicore automatique a base de pthread.
%
%Copperhead
%font du parallelism dans tous les sens :
%\url{http://www.eecs.berkeley.edu/Pubs/TechRpts/2010/EECS-2010-124.pdf}
%mais c'est une syntax particuliére. => ils disent ne aps trop se préoccupé des perfs
%mais plus de la scalabilité
%The runtime operates on CuArrays, which extend Numpy.ndarray
%
%Theano
%C'est un truc fait pour le parallélisme MAIS
%197 * SIMD parallelism on the CPU comes from the compiler.                          
%198 * Multi-core parallelism is only supported by Conv2d(not by default).           
%199   If the external BLAS implementation supports it,                              
%200   there are also, gemm, gemv and ger that are parallelized
%tiré de leurs docs -> D'après le code ils fusionne les opérateur pour que le
%backend puisse essayer de faire du SIMD avec ça.
%Pour le parallel : 
%Multi-core parallelism is only supported by Conv2d(not by default)
%
%Shedskin
%Apparement, pas de SIMD, mais du parallelism explicite a base de Parallel
%Python ou pprocess (fork)
%
%parakeet
%fusion d'opérateur, OpenMP mais pas de SIMD
%\url{http://cs.nyu.edu/web/Research/Theses/eric_h.pdf}
%Attention, ils ont des contraites fortes : si on utilise map, reduce, ..., on
%concidére que c'est paralléle sans vérification. (2.4 de la thèse)
%(D'ailleurs dans Pythran, on vérifie la pureté de la fonction mais on ne
%vérifie pas la sortie :
%ex a = map(foo, xrange(12)); if a == [foo(0), foo(1), ...]: ....
%ne fonctionnerait pas avec l'analyse pmap)
%
%
%Numpypy
%Ils ne font pas ce genre d'optim
%
%
%ATTENTION : le papier risque de faire quand même très répétitif par rapport a WPMVP'14

%TODO ICI UN PARAGRAPH SUR COMMENT LES AUTRES FONT DE la vectorization ET DU PARALLELISME

%TODO un paragraph sur le parallélisme dans numpy

In this papers, we will speak about the Pythran compiler that turn pure Python
code to C++ to create native module and gather high level information to
profiter pleinement de la parallélisation des opérations.

This paper highlight two major contributions. We first introduce how vectoriation
have been added numpy computation intensive kernel thank to
\boostsimd library provided by NT2. We present performances results for Pythran
with and without this auto-vectorization to show performances gains.
Secondly, we introduce parallelization on multi-core architecture for these
computation intensive kernel and evaluates it.

The remainder of this paper is organized as follows. Section \ref{sec:numpy}
describes the differents numpy constructs that are highly parallel and how to
"profiter" from this parallelism. Section \ref{sec:optim} describes how Pythran
optimisation enhance this parallelism increasing computation kernel. Section 
\ref{sec:SIMD} introduces how \boostsimd is used to add auto-vectorization in
Pythran. Section \ref{sec:omp} show use of OpenMP in these kernel for
mutli-core architecture. We will discuss performance result for those
optimization using two different hardware configurations in Section
\ref{sec:perfs}. We finally present some related works in section
\ref{sec:related-work} before concluding.

\section{Data parallelism in Numpy}
\label{sec:numpy}

\section{Optimization feed each other}
\label{sec:optim}

\section{Vectorization using \boostsimd}
\label{sec:SIMD}

\section{Multi core parallelism using OpenMP}
\label{sec:omp}

\section{Benchmarks}
\label{sec:perfs}

\subsection{Pythran with/without Vectorization}

\subsection{Pythran with/without OpenMP}

\subsection{Combining OpenMP and Vectorization}

% Une autre section pour la comparaison avec les autres compilo?

\section{Related works}
\label{sec:related-work}

% On aura pas le temps pour le papier je pense mais faire une comparaison avec du
% C pure pour voir ce que les compilo de C arrive a faire.

\section*{Conclusion}

\acks

This project has been partially funded by the CARP Project and the SILKAN
Company.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{biblio}

\end{document}